---
title: "Final Project"
author: "Nick Gauthier, Daman McBride, Andrew Racek"
date: "May 5, 2026"
format: pdf
execute: 
  echo: false
  warning: false
---

# Plan

The goal of this quarto document is to create a pdf that will have four sections. The first section will intro the data set, describing it case and case attributes and how it meets FAIR and CARE standards. The second section will compare who people said they would vote for before the election and who they said they voted for after the 2024 presidential election with the actual results from the 2024 presidential election. The third section will . The last section will

# Data Set Introduction

The data we will use in our analysis comes from the American National Election Studies (ANES). ANES releases their data every election year, where they ask \~5,000 people a wide range of questions both before and after the election. A case in this data set is a single eligible US voter that took part in this survey. The data set contains \~1,700 case attributes which ask the survey respondent many different questions and will usually give the answer in the form of a number which corresponds to a set of multiple choice answers. Some of the questions are only asked based on previous answers so if the answer is -1 in the data set it is because that question was not asked to that respondent. We each picked a different topic that we wanted to cover using this data set and will elaborate on the specific case attributes we picked in each of our sections. ANES follows the FAIR principles of findable, accessible, interoperable, and reusable as it publicly available data set that can be downloaded straight from their website using a dta file. It has a rich code book that is necessary to see what question each column asks and what number the respondent answered with corresponds to. ANES follows the CARE principles of collective benefit, authority to control, responsibility, and ethics helps people understand how people are feeling on a wide range of topics relating to the election while dealing with being about politics, which is a sensitive issue and people may have privacy concerns over sharing their opinions on this topic.

# Comparison of Election Results to Pre and Post Election Polling

## Plan

The goals of these visuals are to compare the 2024 ANES data to the actual presidential election by comparing the proportion of votes each candidate received in the actual election to both the pre-election polling and the post-election polling.

To do this we need the 2024 ANES data, along with the actual election results, which required both the 2024 Presidential General Election Results and the Total Number of Registered Voters which I got both from two separate tables from the Atlas of US Elections. We also need to library tidyverse, rvest, dplyr, esquisse, plotly, knitr, kableExtra, and haven. Some functions we need include read_dta, read_html, select, mutate, filter, count, recode, rename, full_join, group_by, summarize, bind_cols, pivot_longer, kable, kable_classic, and ggplot.

Steps

1.  Open R
2.  Library haven, tidyverse, rvest, dplyr, esquisse, plotly, knitr, and kableExtra
3.  Use read_dta to store all of the ANES 2024 data
4.  Select just the 8 questions that deal with who people voted for pre and post election along with if they voted early
5.  Mutate the answers from these questions to be numeric
6.  Use select and filter to first only find the answers for whole people voted for if they voted early
7.  Use count to get the count for each possible answer
8.  Mutate to recode the numerical values to the candidate they correspond to
9.  Rename the column with the candidate names from its general name to Candidate
10. Repeat steps 6 through 10, this time selecting and filtering for the pre election polling from non-early voters
11. Use full_join to combine the pre-election polling with the early voters
12. Use summarize and mutate to calculate the proportion of votes each candidate received in the pre election polling and early voter polling
13. Repeat steps 6 through 10 this time selecting and filtering for the post election polling from non-early voters
14. Repeat steps 11 and 12 but with the post election polling from non-early voters with the early voter data to get the proportion of votes by candidate for post election
15. Use read_html and bind_cols to scrape the election results from the US Election Atlas
16. Split this data in two data frames, one that contains the 5 major candidates and the other that contains all other vote receivers
17. Use mutate, group_by, and summarize to combine the data from non-major vote receivers into one row containing the sum of all the non-major votes
18. Use read_html and bind_cols to scrape the number of registered voters data
19. Select the required columns and mutate to make the numbers numeric
20. Use mutate to subtract total votes from total eligible voters to get the number of people who did not vote
21. Use full_join to combine the major candidate votes with the one row of non major votes and the one row of non voters
22. Mutate to find the proportion of votes by candidate
23. Use full_join to combine the pre-election data with the post-election data and the actual election results into one data frame
24. Use mutate to round the proportions and select to reorder the columns
25. To create the table use kable and kable_classic to give the table a title, align each column centrally, set the font size, and alternate the background row color
26. Use pivot_longer to put all of the proportions across each of the three groups into one column where the time of the proportion goes to beforePostReal
27. Create a plot where X is beforePostReal, Y is the proportions, and it is both grouped and colored by candidate. Use both geom_line and geom_point to make points for each proportion and connect them for each of the three times data was collected. Use labs to rename the axes and give the plot a title

## More Information 

The 9 questions I needed to get the voter data from this survey were

1.  V241035, which asks "PRE: ALREADY VOTED IN GENERAL ELECTION"
2.  V241036, which asks "PRE: CONFIRMATION VOTED (EARLY) IN NOVEMBER 5 ELECTION"
3.  V241038, which asks "PRE: DID R VOTE FOR PRESIDENT"
4.  V241039, which asks "PRE: FOR WHOM DID R VOTE FOR PRESIDENT"
5.  V241042, which asks "PRE: DOES R INTEND TO VOTE FOR PRESIDENT"
6.  V241043, which asks "PRE: FOR WHOM DOES R INTEND TO VOTE FOR PRESIDENT"
7.  V242065, which asks "POST: DID R VOTE IN THE 2024 ELECTION"
8.  V242066, which asks "POST: DID R VOTE FOR PRESIDENT"
9.  V242067, which asks "POST: FOR WHOM DID R VOTE FOR PRESIDENT"

The answers to these questions were all in integers, which corresponded to an answer found in the code book. Additional data was need from the US Election Atlas to compare the actual election results. I used two different tables from the US Election Atlas but in both each case represents one eligible US voter. In the Election results table it has columns for the candidate name, the candidate's vice president, the candidate's party, the popular votes received as a number and percentage, and the electoral votes received as a number and percentage. This table shows who eligible voters voted for but we need another table to find the total number of eligible voters to find the number of total eligible voters who did not vote. The second table shows the number of registered voters and number of people who voted by state with columns showing a map of the state, a pie percentage, the state name, the total voting age population, the number of registered voters, the number of votes, the percentage of voting age population that voted, and the percentage of registered voters that voted. The US Election Atlas follows FAIR principles of findable, accessible, interoperable, and reusable as it was easy to access online and the data did not require any initial cleaning, just manipulation to select the values I required. It also follows CARE principles of collective benefit, authority to control, responsibility, and ethics as it showed voter data which is important to help people learn about election results while also protecting the privacy of those who voted by not showing who any individual person voted for. My data analysis was exploratory as I did not know what I wanted to compare when initially looking at the ANES data. I used tidyverse as I needed to mutate the data from the US Election Atlas to compare it to the ANES data.

```{r}
#install.packages("haven")
library("haven")
data2024 <- read_dta("anes_timeseries_2024_stata_20250808 (1).dta")
#data2020 <- read_dta("anes_timeseries_2020_stata_20220210.dta")

library(tidyverse)
library(rvest)
library(dplyr)
library(esquisse)
library(plotly)
library(knitr)
library(kableExtra)


nickData <- data2024 |>
  #Here are the 9 questions I need to get the voter data from this survey
  #V241035 asks "PRE: ALREADY VOTED IN GENERAL ELECTION"
  #V241036 asks "PRE: CONFIRMATION VOTED (EARLY) IN NOVEMBER 5 ELECTION"
  #V241038 asks "PRE: DID R VOTE FOR PRESIDENT"
  #V241039 asks "PRE: FOR WHOM DID R VOTE FOR PRESIDENT"
  #V241042 asks "PRE: DOES R INTEND TO VOTE FOR PRESIDENT"
  #V241043 asks "PRE: FOR WHOM DOES R INTEND TO VOTE FOR PRESIDENT"
  #V242065 asks "POST: DID R VOTE IN THE 2024 ELECTION"
  #V242066 asks "POST: DID R VOTE FOR PRESIDENT"
  #V242067 asks "POST: FOR WHOM DID R VOTE FOR PRESIDENT"
  #Select and mutate these questions to make the answers numeric
  select("V241035", "V241036", "V241038", "V241039", "V241042", "V241043", 
         "V242065", "V242066", "V242067") |>
  mutate(across(c(V241035, V241036, V241038, V241039, V241042, V241043, 
                  V242065, V242066, V242067), as.numeric))


earlyVoters <- nickData |>
  #Filter to those who answered and confirmed they voted early
  filter(V241035 == 1, V241036 == 1) |>
  #Select the question did they vote and if so who did they vote for
  select("V241038", "V241039") |>
  #Remove those who refused to answer who they voted for
  filter(V241039 != -9) |>
  #Select just the who they voted for column
  #Value will be -1 if inapplicable and because of filters only inapplicable if didn't vote
  #Values 1 through 6 are different candidates
  select(V241039) |>
  #Use count to get the number of answers each value has
  count(V241039) |>
  #Mutate to change V241039 to Candidate Names
  mutate(V241039 = recode(V241039,
                          `-1` = "Did Not Vote",
                          `1` = "Kamala D. Harris",
                          `2` = "Donald J. Trump",
                          `3` = "Robert F. Kennedy, Jr.",
                          `4` = "Cornel R. West",
                          `5` = "Jill E. Stein",
                          `6` = "Another Candidate")) |>
  #Use rename to change column name to Candidate
  rename("Candidate" = V241039)


intendedVoters <- nickData |>
  #Filter out those who voted early
  filter(V241035 == 2 | V241036 == 2) |>
  #Select the question will they vote and if so who will they vote for
  select("V241042", "V241043") |>
  #Remove those who refused to answer or didn't know if/who to vote
  filter(V241042 != -9, V241042 != -8, V241043 != -9, V241043 != -8) |>
  #Select just the who they will vote for column
  #Value will be -1 if inapplicable and because of filters only inapplicable if didn't vote
  #Values 1 through 6 are different candidates
  select(V241043) |>
  #Use count to get the number of answers each value has
  count(V241043) |>
  #Mutate to change V241043 to Candidate Names
  mutate(V241043 = recode(V241043,
                          `-1` = "Did Not Vote",
                          `1` = "Kamala D. Harris",
                          `2` = "Donald J. Trump",
                          `3` = "Robert F. Kennedy, Jr.",
                          `4` = "Cornel R. West",
                          `5` = "Jill E. Stein",
                          `6` = "Another Candidate")) |>
  #Use rename to change column name to Candidate
  rename("Candidate" = V241043)


preElectionData <- intendedVoters |>
  #Use full_join to combine early voter data to show all pre election data
  full_join(earlyVoters) |>
  #group by candidate and summarize n as sum of n to add the totals together
  group_by(Candidate) |>
  summarize(n = sum(n)) |>
  #Mutate to create new column seeing the proportion each candidate got
  mutate("Pre Election Proportion" = n / sum(n)) |>
  rename("Pre Election Voters" = n)


postElectionVoters <- nickData |>
  #Filter out those who voted early
  #Do this because they weren't asked these questions in post
  filter(V241035 == 2 | V241036 == 2) |>
  #Select the question did they vote, did they vote for president and if so who did they vote for
  select(V242065, V242066, V242067) |>
  #Filter out those who did not give an answer to whether they voted
  filter(V242065 != -9, V242065 != -8, V242065 != -7, V242065 != -6) |>
  #Select the did they vote for president and who they voted for president columns
  select(V242066, V242067) |>
  #Filter out those who did not give an answer for if/who they voted 
  filter(V242066 != -1, V242067 != -9, V242067 != -8) |>
  #Select just the who they voted for column
  #Value will be -1 if inapplicable and because of filters only inapplicable if didn't vote
  #Values 1 through 6 are different candidates
  select(V242067) |>
  #Use count to get the number of answers each value has
  count(V242067) |>
  #Mutate to change V242067 to Candidate Names
  mutate(V242067 = recode(V242067,
                          `-1` = "Did Not Vote",
                          `1` = "Kamala D. Harris",
                          `2` = "Donald J. Trump",
                          `4` = "Cornel R. West",
                          `5` = "Jill E. Stein",
                          `6` = "Another Candidate")) |>
  #Use rename to change column names to Candidate and Post Election Voters
  rename("Candidate" = V242067)


postElectionData <- postElectionVoters |>
  #Use full_join to combine early voter data to show all votes
  full_join(earlyVoters) |>
  #group by candidate and summarize n as sum of n to add the totals together
  group_by(Candidate) |>
  summarize(n = sum(n)) |>
  #Mutate to create new column seeing the proportion each candidate got
  mutate("Post Election Proportion" = n / sum(n)) |>
  rename("Post Election Voters" = n)


#read_html voting result
votingResultList <- read_html(
  x = "https://uselectionatlas.org/RESULTS/national.php?year=2024&minper=0&off=0&elect=0"
) |>
  html_elements(css = "table") |>
  html_table()

#bind_cols to turn votingResultList into a data frame votingResultRaw
votingResultRaw <- bind_cols(votingResultList[[2]])

#Remove first 4 rows and all rows after 30
votingResult <- votingResultRaw[c(5:30),] |>
  #Rename and select the columns required
  rename("Candidate" = 2, "Election Voters" = 5) |>
  select(Candidate, "Election Voters") |>
  #Remove commas to convert to numeric
  mutate(across("Election Voters", ~ gsub(",", "", .))) |>
  mutate(across("Election Voters", as.numeric))

votingMajorCandidates <- votingResult |>
  #Filter out minor candidates
  filter(Candidate == "Donald J. Trump" |
           Candidate == "Kamala D. Harris" |
           Candidate == "Robert F. Kennedy, Jr." |
           Candidate == "Cornel R. West" |
           Candidate == "Jill E. Stein")

votingMinorCandidates <- votingResult |>
  #Remove major candidates
  filter(Candidate != "Donald J. Trump" &
           Candidate != "Kamala D. Harris" &
           Candidate != "Robert F. Kennedy, Jr." &
           Candidate != "Cornel R. West" &
           Candidate != "Jill E. Stein") |>
  #Rename all renaming Candidates as same name to group
  mutate(Candidate = "Another Candidate") |>
  #Group by candidate and summarize Election voters to get one row 
  #with sum of all votes from remaining candidates
  group_by(Candidate) |>
  summarize(`Election Voters` = sum(`Election Voters`))


#read_html voter turnout
voterTurnoutList <- read_html(
  x = "https://uselectionatlas.org/RESULTS/data.php?year=2024&datatype=national&def=vto&f=0&off=0&elect=0"
) |>
  html_elements(css = "table") |>
  html_table()

#bind_cols to turn voterTurnoutList into a data frame voterTurnoutRaw
voterTurnoutRaw <- bind_cols(voterTurnoutList[[3]])

#Remove first 2 rows
voterTurnout <- voterTurnoutRaw[-c(1:2),] |>
  #Rename and select the columns required
  rename("Candidate" = 3, "Eligible Voters" = 4, "Votes" = 6) |>
  select(Candidate, "Eligible Voters", Votes) |>
  #Filter candidate to get the required row
  filter(Candidate  == "Total") |>
  #Remove commas to convert to numeric
  mutate(across(c("Eligible Voters", Votes), ~ gsub(",", "", .))) |>
  mutate(across(c("Eligible Voters", Votes), as.numeric)) |>
  #Create new column Election Voters 
  #where the row has the value of the number of people who didn't vote
  mutate(`Election Voters` = `Eligible Voters` - Votes) |>
  #Mutate to change Total to Did Not Vote
  mutate(Candidate = recode(Candidate, "Total" = "Did Not Vote")) |>
  #Select the 2 required columns for the final table
  select(Candidate, "Election Voters")


electionData <- votingMajorCandidates |>
  #Use full_join to bring all election data into one data frame
  full_join(votingMinorCandidates) |>
  full_join(voterTurnout) |>
  #Create proportion column by dividing election voters by sum of all election voters
  mutate("Election Proportion" = `Election Voters` / sum(`Election Voters`))


allElectionData <- electionData |>
  #Use full join by Candidate to combine all data into one data frame
  full_join(postElectionData, by = "Candidate") |>
  full_join(preElectionData, by = "Candidate") |>
  #Round the proportions to 4 digits
  mutate(across("Post Election Proportion", function(x) round(x, 4))) |>
  mutate(across("Pre Election Proportion", function(x) round(x, 4))) |>
  mutate(across("Election Proportion", function(x) round(x, 4))) |>
  #Use select to reorder columns
  select(Candidate, "Pre Election Voters", "Pre Election Proportion",
         "Post Election Voters", "Post Election Proportion",
         "Election Voters", "Election Proportion")


#Use kable and kable_classic to title, align, change font, and alternate row color
allElectionData |>
  #Add commas back into large numbers
  mutate(across(where(is.numeric), ~ format(.x, big.mark = ","))) |>
  kable(
    caption = "Votes and Proportion of Votes by Candidate",
    align = "c",
    format = "latex"
  ) |>
  kable_classic(
    font_size = 12,
    lightable_options = "striped"
  ) |>
  kable_styling(latex_options = c("scale_down", "hold_position"))
```

## Votes and Proportion of Votes by Candidate Table Analysis

The initial table shows the number of votes each candidate received along with the proportion of the total votes they received. From the polling data we can see that Kamala D. Harris had the highest proportion in the polling data from both the pre and post election polling yet in the actual election data Donald J. Trump received another 2.3 million votes, giving him a slightly higher proportion than Kamala D. Harris. The biggest change from the survey data to the actual election was that not voting went from a small proportion in the polling data to the largest proportion in the actual election

```{r}
electionDataPlot <- allElectionData |>
  #rename the columns to make them better names for plot
  rename("Before" = "Pre Election Proportion", 
         "Post" = "Post Election Proportion",
         "Real" = "Election Proportion") |>
  #pivot longer to have all proportions in one column 
  #where the timing of the proportion is in another column
  pivot_longer(cols = c("Before", "Post", "Real"),
               names_to = "beforePostReal", 
               values_to = "Proportion") 


#Create plot where x is the timing, y is the proportion, and it is grouped by candidate
ggplot(electionDataPlot, aes(x = beforePostReal, y = Proportion, group = Candidate)) +
  #Use point to plot the proportions
  geom_point(aes(color = Candidate)) +
  #Use line to connect the points
  geom_line(aes(color = Candidate)) +
  #Use labs to give the plot a title, and rename the axes
  labs(
    title = "Proportion of Votes by Candidate",
    x = "Before Election vs Post Election vs Actual Election",
    y = "Proportion of Votes",
    color = "Candidate") +
  theme_bw()
```

## Proportion of Votes by Candidate Plot Description 

Plot has a y value of the proportion of votes a candidate received where the x value is three different times these proportions were gathered. These points are then connected to compare the proportion of votes a candidate received across the three polling times. The plot shows that the number of people who did not vote in the polling data was a very small proportion but that proportion rose dramatically in the actual election. Polling data also had Kamala D. Harris have the highest proportion in both pre and post election data but in the actual election the highest proportion was from eligible voters not voting, following by Donald J. Trump.

## Analysis of Both Visualizations

When first looking at both data visualizations the clear change in proportion is seen between not voting in the surveys to not voting in the actual election. From this we learn that a lot of people when surveyed said they were going to vote for a candidate in both the pre and post election surveys but when looking at the actual election it seems that they were not motivated enough to actually go out and vote. This shows that the current methods of polling people must find additional ways to try and survey the likelihood of people actually going out and voting because when people are surveyed their answers may change from what they actually may end up voting. This is also seen in this data as in the polling Kamala D. Harris has a slight lead over Donald J. Trump but in the actual election Donald J. Trump ended up getting 2.3 million more votes. To account for the fact that people may answer differently in polls than what they may actually end up voting, surveyors must come up with new unique ways to try and get the actual answer from the people, which if I were to survey people about election data I would ask them questions like who they think will win or who they think people around them might be voting for and then comparing those to the election results to see if we get more accurate data.

# (Title for Analysis)

## Plan

The goals of these visuals are to . Needs. Steps

```{r}
```

## Table 1 Analysis

The initial table shows .

```{r}
```

## Plot 2 Description and Analysis of Both Visualizations

Plot has .

When first looking at both data visualizations.

# (Title for Analysis)

## Plan

The goals of these visuals are to . Needs. Steps.

```{r}
```

## Table 1 Analysis

The initial table shows .

```{r}
```

## Plot 2 Description and Analysis of Both Visualizations

Plot has .

When first looking at both data visualizations.

# Appendix 1

## Proportion of Votes by Candidates code to wrangle and create plots by Nick Gauthier

#install.packages("haven") library("haven") data2024 \<- read_dta("anes_timeseries_2024_stata_20250808 (1).dta") #data2020 \<- read_dta("anes_timeseries_2020_stata_20220210.dta")

library(tidyverse) library(rvest) library(dplyr) library(esquisse) library(plotly) library(knitr) library(kableExtra)

nickData \<- data2024 \|\> #Here are the 9 questions I need to get the voter data from this survey #V241035 asks "PRE: ALREADY VOTED IN GENERAL ELECTION" #V241036 asks "PRE: CONFIRMATION VOTED (EARLY) IN NOVEMBER 5 ELECTION" #V241038 asks "PRE: DID R VOTE FOR PRESIDENT" #V241039 asks "PRE: FOR WHOM DID R VOTE FOR PRESIDENT" #V241042 asks "PRE: DOES R INTEND TO VOTE FOR PRESIDENT" #V241043 asks "PRE: FOR WHOM DOES R INTEND TO VOTE FOR PRESIDENT" #V242065 asks "POST: DID R VOTE IN THE 2024 ELECTION" #V242066 asks "POST: DID R VOTE FOR PRESIDENT" #V242067 asks "POST: FOR WHOM DID R VOTE FOR PRESIDENT" #Select and mutate these questions to make the answers numeric select("V241035", "V241036", "V241038", "V241039", "V241042", "V241043", "V242065", "V242066", "V242067") \|\> mutate(across(c(V241035, V241036, V241038, V241039, V241042, V241043, V242065, V242066, V242067), as.numeric))

earlyVoters \<- nickData \|\> #Filter to those who answered and confirmed they voted early filter(V241035 == 1, V241036 == 1) \|\> #Select the question did they vote and if so who did they vote for select("V241038", "V241039") \|\> #Remove those who refused to answer who they voted for filter(V241039 != -9) \|\> #Select just the who they voted for column #Value will be -1 if inapplicable and because of filters only inapplicable if didn't vote #Values 1 through 6 are different candidates select(V241039) \|\> #Use count to get the number of answers each value has count(V241039) \|\> #Mutate to change V241039 to Candidate Names mutate(V241039 = recode(V241039, `-1` = "Did Not Vote", `1` = "Kamala D. Harris", `2` = "Donald J. Trump", `3` = "Robert F. Kennedy, Jr.", `4` = "Cornel R. West", `5` = "Jill E. Stein", `6` = "Another Candidate")) \|\> #Use rename to change column name to Candidate rename("Candidate" = V241039)

intendedVoters \<- nickData \|\> #Filter out those who voted early filter(V241035 == 2 \| V241036 == 2) \|\> #Select the question will they vote and if so who will they vote for select("V241042", "V241043") \|\> #Remove those who refused to answer or didn't know if/who to vote filter(V241042 != -9, V241042 != -8, V241043 != -9, V241043 != -8) \|\> #Select just the who they will vote for column #Value will be -1 if inapplicable and because of filters only inapplicable if didn't vote #Values 1 through 6 are different candidates select(V241043) \|\> #Use count to get the number of answers each value has count(V241043) \|\> #Mutate to change V241043 to Candidate Names mutate(V241043 = recode(V241043, `-1` = "Did Not Vote", `1` = "Kamala D. Harris", `2` = "Donald J. Trump", `3` = "Robert F. Kennedy, Jr.", `4` = "Cornel R. West", `5` = "Jill E. Stein", `6` = "Another Candidate")) \|\> #Use rename to change column name to Candidate rename("Candidate" = V241043)

preElectionData \<- intendedVoters \|\> #Use full_join to combine early voter data to show all pre election data full_join(earlyVoters) \|\> #group by candidate and summarize n as sum of n to add the totals together group_by(Candidate) \|\> summarize(n = sum(n)) \|\> #Mutate to create new column seeing the proportion each candidate got mutate("Pre Election Proportion" = n / sum(n)) \|\> rename("Pre Election Voters" = n)

postElectionVoters \<- nickData \|\> #Filter out those who voted early #Do this because they weren't asked these questions in post filter(V241035 == 2 \| V241036 == 2) \|\> #Select the question did they vote, did they vote for president and if so who did they vote for select(V242065, V242066, V242067) \|\> #Filter out those who did not give an answer to whether they voted filter(V242065 != -9, V242065 != -8, V242065 != -7, V242065 != -6) \|\> #Select the did they vote for president and who they voted for president columns select(V242066, V242067) \|\> #Filter out those who did not give an answer for if/who they voted filter(V242066 != -1, V242067 != -9, V242067 != -8) \|\> #Select just the who they voted for column #Value will be -1 if inapplicable and because of filters only inapplicable if didn't vote #Values 1 through 6 are different candidates select(V242067) \|\> #Use count to get the number of answers each value has count(V242067) \|\> #Mutate to change V242067 to Candidate Names mutate(V242067 = recode(V242067, `-1` = "Did Not Vote", `1` = "Kamala D. Harris", `2` = "Donald J. Trump", `4` = "Cornel R. West", `5` = "Jill E. Stein", `6` = "Another Candidate")) \|\> #Use rename to change column names to Candidate and Post Election Voters rename("Candidate" = V242067)

postElectionData \<- postElectionVoters \|\> #Use full_join to combine early voter data to show all votes full_join(earlyVoters) \|\> #group by candidate and summarize n as sum of n to add the totals together group_by(Candidate) \|\> summarize(n = sum(n)) \|\> #Mutate to create new column seeing the proportion each candidate got mutate("Post Election Proportion" = n / sum(n)) \|\> rename("Post Election Voters" = n)

#read_html voting result votingResultList \<- read_html( x = "[https://uselectionatlas.org/RESULTS/national.php?year=2024&minper=0&off=0&elect=0"](https://uselectionatlas.org/RESULTS/national.php?year=2024&minper=0&off=0&elect=0") ) \|\> html_elements(css = "table") \|\> html_table()

#bind_cols to turn votingResultList into a data frame votingResultRaw votingResultRaw \<- bind_cols(votingResultList\[\[2\]\])

#Remove first 4 rows and all rows after 30 votingResult \<- votingResultRaw\[c(5:30),\] \|\> #Rename and select the columns required rename("Candidate" = 2, "Election Voters" = 5) \|\> select(Candidate, "Election Voters") \|\> #Remove commas to convert to numeric mutate(across("Election Voters", \~ gsub(",", "", .))) \|\> mutate(across("Election Voters", as.numeric))

votingMajorCandidates \<- votingResult \|\> #Filter out minor candidates filter(Candidate == "Donald J. Trump" \| Candidate == "Kamala D. Harris" \| Candidate == "Robert F. Kennedy, Jr." \| Candidate == "Cornel R. West" \| Candidate == "Jill E. Stein")

votingMinorCandidates \<- votingResult \|\> #Remove major candidates filter(Candidate != "Donald J. Trump" & Candidate != "Kamala D. Harris" & Candidate != "Robert F. Kennedy, Jr." & Candidate != "Cornel R. West" & Candidate != "Jill E. Stein") \|\> #Rename all renaming Candidates as same name to group mutate(Candidate = "Another Candidate") \|\> #Group by candidate and summarize Election voters to get one row #with sum of all votes from remaining candidates group_by(Candidate) \|\> summarize(`Election Voters` = sum(`Election Voters`))

#read_html voter turnout voterTurnoutList \<- read_html( x = "[https://uselectionatlas.org/RESULTS/data.php?year=2024&datatype=national&def=vto&f=0&off=0&elect=0"](https://uselectionatlas.org/RESULTS/data.php?year=2024&datatype=national&def=vto&f=0&off=0&elect=0") ) \|\> html_elements(css = "table") \|\> html_table()

#bind_cols to turn voterTurnoutList into a data frame voterTurnoutRaw voterTurnoutRaw \<- bind_cols(voterTurnoutList\[\[3\]\])

#Remove first 2 rows voterTurnout \<- voterTurnoutRaw\[-c(1:2),\] \|\> #Rename and select the columns required rename("Candidate" = 3, "Eligible Voters" = 4, "Votes" = 6) \|\> select(Candidate, "Eligible Voters", Votes) \|\> #Filter candidate to get the required row filter(Candidate == "Total") \|\> #Remove commas to convert to numeric mutate(across(c("Eligible Voters", Votes), \~ gsub(",", "", .))) \|\> mutate(across(c("Eligible Voters", Votes), as.numeric)) \|\> #Create new column Election Voters #where the row has the value of the number of people who didn't vote mutate(`Election Voters` = `Eligible Voters` - Votes) \|\> #Mutate to change Total to Did Not Vote mutate(Candidate = recode(Candidate, "Total" = "Did Not Vote")) \|\> #Select the 2 required columns for the final table select(Candidate, "Election Voters")

electionData \<- votingMajorCandidates \|\> #Use full_join to bring all election data into one data frame full_join(votingMinorCandidates) \|\> full_join(voterTurnout) \|\> #Create proportion column by dividing election voters by sum of all election voters mutate("Election Proportion" = `Election Voters` / sum(`Election Voters`))

allElectionData \<- electionData \|\> #Use full join by Candidate to combine all data into one data frame full_join(postElectionData, by = "Candidate") \|\> full_join(preElectionData, by = "Candidate") \|\> #Round the proportions to 4 digits mutate(across("Post Election Proportion", function(x) round(x, 4))) \|\> mutate(across("Pre Election Proportion", function(x) round(x, 4))) \|\> mutate(across("Election Proportion", function(x) round(x, 4))) \|\> #Use select to reorder columns select(Candidate, "Pre Election Voters", "Pre Election Proportion", "Post Election Voters", "Post Election Proportion", "Election Voters", "Election Proportion")

#Use kable and kable_classic to title, align, change font, and alternate row color allElectionData \|\> #Add commas back into large numbers mutate(across(where(is.numeric), \~ format(.x, big.mark = ","))) \|\> kable( caption = "Votes and Proportion of Votes by Candidate", align = "c", format = "latex" ) \|\> kable_classic( font_size = 18, lightable_options = "striped" )\|\> kable_styling(latex_options = c("scale_down", "hold_position"))

electionDataPlot \<- allElectionData \|\> #rename the columns to make them better names for plot rename("Before" = "Pre Election Proportion", "Post" = "Post Election Proportion", "Real" = "Election Proportion") \|\> #pivot longer to have all proportions in one column #where the timing of the proportion is in another column pivot_longer(cols = c("Before", "Post", "Real"), names_to = "beforePostReal", values_to = "Proportion")

#Create plot where x is the timing, y is the proportion, and it is grouped by candidate ggplot(electionDataPlot, aes(x = beforePostReal, y = Proportion, group = Candidate)) + #Use point to plot the proportions geom_point(aes(color = Candidate)) + #Use line to connect the points geom_line(aes(color = Candidate)) + #Use labs to give the plot a title, and rename the axes labs( title = "Proportion of Votes by Candidate", x = "Before Election vs Post Election vs Actual Election", y = "Proportion of Votes", color = "Candidate") + theme_bw()

## Code to wrangle and generate the plots

code

## Code to wrangle and generate the plots

code
